---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(actvotes)
library(dplyr)
library(ggplot2)

```

# actvotes

<!-- badges: start -->
[![R build status](https://github.com/bryceroney/actvotes/workflows/R-CMD-check/badge.svg)](https://github.com/bryceroney/actvotes/actions)
<!-- badges: end -->

This package provides tidy datasets of ACT territory election results.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bryceroney/actvotes")
```
## Datasets
This package currently contains the following datasets from the [ACT Electoral Commission](https://www.elections.act.gov.au):

* 2020 Election: Polling Places (`act_polling_places_2020`), Candidates (`act_candidates_2020`), First Preference (`act_fp_2020`) results and final preference distribution (`act_preferences_2020`).

## Usage Example

```{r plot, fig.height=6, fig.width=7, dpi=220}

act_fp_2020 %>%
  inner_join(act_candidates_2020, by=c("candidate"="ballot_paper_name")) %>%
  filter(polling_place == 'Manuka' & electorate == 'Kurrajong' &
           party %in% c('ACT Labor', 'Canberra Liberals', 'The ACT Greens')) %>%
  mutate(candidate = reorder(candidate, votes),
         votes = votes/sum(votes)) %>%
  ggplot(aes(x=candidate, y=votes)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~party, scales='free_y', ncol=1) +
  labs(
    title = 'Votes by candidate for the three major parties',
    subtitle = 'Manuka polling place',
    y = 'Votes (%)',
    x = NULL,
    caption = 'Source: ACT Electoral Commission'
  ) +
  theme_bw()
```
