---
title: "Party Primary Vote Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Party Primary Vote Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(actvotes)
```

## Graphing election results
The below graph shows each of the three major parties' strongest booths across the ACT:

```{r graphs, fig.width=6, fig.height=4}
library(dplyr)
library(ggplot2)

act_fp_2020 %>%
  inner_join(act_candidates_2020, by=c("candidate"="ballot_paper_name")) %>%
  group_by(polling_place, party) %>%
  summarise(votes = sum(votes)) %>%
  add_tally(wt=votes) %>%
  mutate(percent = votes/n) %>%
  ungroup() %>%
  filter(party %in% c('ACT Labor', 'Canberra Liberals', 'The ACT Greens')) %>%
  group_by(party) %>%
  top_n(5, wt=percent) %>%
  mutate(polling_place = reorder(polling_place, percent)) %>%
  ggplot(aes(x=polling_place, y=percent)) +
  geom_col() +
  facet_wrap(~party, scales='free_y', ncol=1) +
  coord_flip() +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(x=NULL, y='First Preferences (%)') +
  theme_minimal()
```

## Mapping results
The `act_polling_places_2020` object is an `sf` object which has spatial data for each of the polling places.

The below example is a simple `leaflet::leaflet()` map, with the size of the circles determined by the total number of votes cast.

You can do other analyses, including in `ggplot2` with `ggplot2::geom_sf()`

```{r map, fig.width=6, warning=FALSE}
library(sf)
library(leaflet)

votes_by_polling_place <- act_fp_2020 %>%
  group_by(polling_place) %>%
  summarise(votes=sum(votes)) %>%
  ungroup() 

act_polling_places_2020 %>%
  inner_join(votes_by_polling_place, by=c("Name"="polling_place")) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(label = ~paste(Name, '- Votes:', votes), radius=~votes/100)

```
